---
title: "Stats_Project"
author: "Grace Huang"
date: "2024-04-01"
output: html_document
---

Add the Libraries
```{r}
library(pastecs)
library(car)
library(QuantPsyc)
library(ggplot2)
library(psych)
library(stats)
```


Import CSV file
```{r}

# Read WPM data
WPM <- read.csv('Raw_WPM.csv', sep = ',', header = TRUE)

# Read Accuracy data
Accuracy <- read.csv('Raw_Accuracy.csv', sep = ',', header = TRUE)

```


# 1. Summary of data

```{r}
# Summary of WPM data
summary(WPM)

# Summary of Accuracy data
summary(Accuracy)

```

# 2. Create Boxplots

```{r}

# Create boxplots for WPM
boxplot(WPM[, c("Silence", "Classical", "Pop")], 
        main = "Words Per Minute by Genre",
        ylab = "Words Per Minute",
        col = c("blue", "green", "red"),
        names = c("Silence", "Classical", "Pop"))

# Create boxplots for Accuracy
boxplot(Accuracy[, c("Silence", "Classical", "Pop")], 
        main = "Accuracy by Genre",
        ylab = "Accuracy",
        col = c("blue", "green", "red"),
        names = c("Silence", "Classical", "Pop"))


# There is a potential outlier in pop with a very low value but after outlier analysis we see that it's not
```


# 3. Pairwise T-test

```{r}
##########
# WPM
##########

# Compare Silence to Classical
t.test(WPM$Silence, WPM$Classical, paired = TRUE)

# Compare Classical to Pop
t.test(WPM$Classical, WPM$Pop, paired = TRUE)

# Compare Silence to Pop
t.test(WPM$Silence, WPM$Pop, paired = TRUE)

```

```{r}

##############
# ACCURACY
##############

# Compare Silence to Classical
t.test(Accuracy$Silence, Accuracy$Classical, paired = TRUE)

# Compare Classical to Pop
t.test(Accuracy$Classical, Accuracy$Pop, paired = TRUE)

# Compare Silence to Pop
t.test(Accuracy$Silence, Accuracy$Pop, paired = TRUE)


```




# 4. One-Factor ANOVA
Anova checks for sphericity
sphereicity is that the paired different samples all have the same variance
```{r}

WPM2 <- read.csv('WPM_2.csv', sep = ',', header = TRUE)

res.aov_WPM <- aov(WPM ~ Condition+User, data = WPM2)

summary(res.aov_WPM)

```


# 5. Scatterplot checks for outliers and independence

```{r}

WPM_Silence <- read.csv('WPM_Silence.csv', sep = ',', header = TRUE)

Silence_scatter <- ggplot(WPM_Silence, aes(x = User, y = WPM)) + geom_point() + labs(x="User", y="Words Per Minute") + ggtitle("User vs Silence WPM Scatterplot")
Silence_scatter


WPM_Classical <- read.csv('WPM_Classical.csv', sep = ',', header = TRUE)

Classical_scatter <- ggplot(WPM_Classical, aes(x = User, y = WPM)) + geom_point() + labs(x="User", y="Words Per Minute") + ggtitle("User vs Classical WPM Scatterplot")
Classical_scatter

WPM_Pop <- read.csv('WPM_Pop.csv', sep = ',', header = TRUE)
Pop_scatter <- ggplot(WPM_Pop, aes(x = User, y = WPM)) + geom_point() + labs(x="User", y="Words Per Minute") + ggtitle("User vs Pop WPM Scatterplot")
Pop_scatter

```



# 6. Check all assumptions of ANOVA Table (Normally distributed)
```{r}

# SILENCE
Silence_hist <- ggplot(WPM, aes(Silence)) + geom_histogram(binwidth = 5, color = "black", fill = "blue") + labs(x = "Silence", y = "Frequency") + ggtitle("Silence Histogram")

Silence_hist

# CLASSICAL
Classical_hist <- ggplot(WPM, aes(Classical)) + geom_histogram(binwidth = 5, color = "black", fill = "blue") + labs(x = "Classical", y = "Frequency") + ggtitle("Classical Histogram")

Classical_hist

# POP
Pop_hist <- ggplot(WPM, aes(Pop)) + geom_histogram(binwidth = 5, color = "black", fill = "blue") + labs(x = "Pop", y = "Frequency") + ggtitle("Pop Histogram")

Pop_hist
```


# 7. Sphericity Test
```{r}

Sphericity <- WPM %>%
    transmute(
      'Classical - Silence' = Classical - Silence,
      'Classical - Pop' = Classical - Pop,
      'Silence - Pop' = Silence - Pop,
    )
Sphericity %>% map(var)
```

# 8. SHAPRIO WILKS TEST
```{r}

shapiro.test(WPM$Silence)
shapiro.test(WPM$Classical)
shapiro.test(WPM$Pop)

# Since the P-values of the Shapiro test of all three categories are greater than 0.05, we can reject the null hypothesis that "Silence", "Classical", and "Pop" are not normally distribted. Therefore, we can say that our data values are likely normally distributed
```




# 9. Use QQ Plot
```{r}
Silence_lm <- lm(WPM ~ User, data = WPM_Silence)
Silence_resid <- resid(Silence_lm)

Silence.student <- rstudent(Silence_lm)
Silence.standard <- rstandard(Silence_lm)

plot(Silence_lm, which = 2, main = "Silence Q-Q Plot of Residuals")

Silence.student
Silence.standard

```

```{r}

Classical_lm <- lm(WPM ~ User, data = WPM_Classical)
Classical_resid <- resid(Classical_lm)

Classical.student <- rstudent(Classical_lm)
Classical.standard <- rstandard(Classical_lm)

plot(Classical_lm, which = 2, main = "Classical Q-Q Plot of Residuals")

Classical.student
Classical.standard


```

```{r}

Pop_lm <- lm(WPM ~ User, data = WPM_Pop)
Pop_resid <- resid(Pop_lm)

Pop.student <- rstudent(Pop_lm)
Pop.standard <- rstandard(Pop_lm)

plot(Pop_lm, which = 2, main = "Pop Q-Q Plot of Residuals")

Pop.student
Pop.standard



# leverage is how much power each point has to influence the slope
```



# 10. Residual Plot
```{r}
# Silence Standardized and Studentized Plots
ols_plot_resid_stand(Silence_lm)
ols_plot_resid_stud(Silence_lm)

```
```{r}
# Classical Standardized and Studentized Plots
ols_plot_resid_stand(Classical_lm)
ols_plot_resid_stud(Classical_lm)
```

```{r}
# Pop Standardized and Studentized Plots
ols_plot_resid_stand(Pop_lm)
ols_plot_resid_stud(Pop_lm)
```
